Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-15 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
before editing sect. 5

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-examples-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$L = ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that:
\begin{enumerate}
  \item $L$ is strongly linearly separable with margin $\gamma$,
  \item $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$,
\end{enumerate}
then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

%$A
%= \sum_{t ~:~ S_t \neq \emptyset} z_t$
%B = \sum_{t ~:~ S_t = \emptyset} z_t
%\sum_{t ~:~ S_t = \emptyset}
\begin{proof}
Let $M = \sum_{t=1}^T z_t$ be the number of
mistakes Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} makes.
Let $A = \sum_{t=1}^T \one[S_t \neq \emptyset] z_t$ be the number of mistakes in the rounds
when $S_t \neq \emptyset$, i.e. the number of rounds line~\ref{line:neg-update} is
executed.
In addition, let $B = \sum_{t=1}^T \one[S_t = \emptyset] z_t$ be the
number of mistakes in the rounds when $S_t = \emptyset$.
It can be easily seen that $M = A + B$.

@@

>>>>

\begin{theorem}[Mistake upper bound]
\label{theorem:strongly-separable-examples-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$L = ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that:
\begin{enumerate}
  \item $L$ is strongly linearly separable with margin $\gamma$,
  \item for all $t$ in $\cbr{1,\ldots,T}$, $\norm{x_t} \le R$,
\end{enumerate}
then the
expected number of mistakes
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

%$\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
%$A
%= \sum_{t ~:~ S_t \neq \emptyset} z_t$
%B = \sum_{t ~:~ S_t = \emptyset} z_t
%\sum_{t ~:~ S_t = \emptyset}
\begin{proof}
Let $M = \sum_{t=1}^T z_t$ be the number of
mistakes Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} makes.
Let $A = \sum_{t=1}^T \one[S_t \neq \emptyset] z_t$ be the number of mistakes in the rounds
when $S_t \neq \emptyset$, i.e. the number of rounds line~\ref{line:neg-update} is
executed.
In addition, let $B = \sum_{t=1}^T \one[S_t = \emptyset] z_t$ be the
number of mistakes in the rounds when $S_t = \emptyset$.
It can be easily seen that $M = A + B$.

@@

The following theorem is a direct consequence of Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound}.

\begin{theorem}[Mistake upper bound for kernelized algorithm]
\label{theorem:kernelized-upper-bound}
Let $k(\cdot, \cdot)$ be a kernel function with feature map $\phi$, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$L = ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that:
\begin{enumerate}
  \item $((\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T))$ is strongly linearly separable with margin $\gamma$,
  \item for all $t$ in $\cbr{1,\ldots,T}$, $k(x_t,x_t) \le R^2$,
\end{enumerate}
then the
expected number of mistakes
Algorithm~\ref{algorithm:kernelized}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

### 2019-01-15 ICML-submission/from-weak-separability-to-strong-separability.tex
Small fixes
before editing sect. 5

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
\begin{align*}
p(x)
& = p(x_1, x_2, \dots, x_d) \\
& = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\end{align*}
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$ is a real coefficient. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
The \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as a
Euclidean norm of the vector of the coefficients of the polynomial.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$.
Let
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(2/\gamma) \right \rceil \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
Then, the labeled examples $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$ are strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$ and $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

@@

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors $(w_i - w_j)/2$ for
$j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors have norm at most
$1$.
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
imply that there exists multivariate polynomials $p_i:\R^d \to \R$ and $q_i:\R^d
\to \R$ such that
\begin{align*}
\deg(p_i) & = \lceil \log_2(2K-2) \rceil \cdot \left\lceil \sqrt{\frac{2}{\gamma}} \right\rceil \; , \\
\deg(q_i) & = (2s+1) r(K-1) \; .
\end{align*}
Futhermore, for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$,
$q_i(x_t) \ge \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and if $y_t \neq i$ then
$p_i(x_t) \le -1/2$, $q_i(x) \le - \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and
\begin{align*}
\norm{p_i} & \le \left(188 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\norm{q_i} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
By \autoref{lemma:norm-bound} there exists $c_i, c_i' \in \ell_2$ such that
$\ip{c_i}{\phi(x)} = p_i(x)$ and $\ip{c_i'}{\phi(x)} = q_i(x)$ and
\begin{align*}
\norm{c_i}_{\ell_2}
& \le \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\norm{c'_i}_{\ell_2} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
Define vectors $u_i, u_i' \in \ell_2$ as
\begin{align*}
u_i & = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; . \\
u_i' & = \frac{c_i'  \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)}}{(2K-5/2) 2^{K-1}} \; .
\end{align*}

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
$\norm{u_1'}^2 + \norm{u_2'}^2 + \dots + \norm{u_K'}^2 \le 1$.
Futhermore, for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma_1$ and
$\ip{u'_{y_t}}{x_t} \ge \gamma_2$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma_1$ and $\ip{u'_j}{x_t} \le - \gamma_2$. In other words,
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_1$ and also strongly linearly
separable with margin $\gamma_2$. Therefore, the examples are strongly separable
with margin $\gamma' = \max\{\gamma_1, \gamma_2\}$.
\end{proof}

>>>>

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
\begin{align*}
p(x)
& = p(x_1, x_2, \dots, x_d) \\
& = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\end{align*}
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$'s are real coefficients. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
Following the terminology of~\cite{Klivans-Servedio-2008},
the \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as the
Euclidean norm of the vector of the coefficients of the polynomial.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$.
Let
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(2/\gamma) \right \rceil \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
Then, the labeled examples $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$ are strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$ and $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

@@

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors $(w_i - w_j)/2$ for
$j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors have norm at most
$1$.
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
imply that there exists multivariate polynomials $p_i:\R^d \to \R$ and $q_i:\R^d
\to \R$ such that
\begin{align*}
\deg(p_i) & = \lceil \log_2(2K-2) \rceil \cdot \left\lceil \sqrt{\frac{2}{\gamma}} \right\rceil \; , \\
\deg(q_i) & = (2s+1) r(K-1) \; .
\end{align*}
Futhermore, for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$,
$q_i(x_t) \ge \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and if $y_t \neq i$ then
$p_i(x_t) \le -1/2$, $q_i(x) \le - \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and
\begin{align*}
\norm{p_i} & \le \left(188 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\norm{q_i} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
By \autoref{lemma:norm-bound} there exists $c_i, c_i' \in \ell_2$ such that
$\ip{c_i}{\phi(x)} = p_i(x)$ and $\ip{c_i'}{\phi(x)} = q_i(x)$ and
\begin{align*}
\norm{c_i}_{\ell_2}
& \le \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\norm{c'_i}_{\ell_2} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
Define vectors $u_i, u_i' \in \ell_2$ as
\begin{align*}
u_i & = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; . \\
u_i' & = \frac{c_i'  \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)}}{\sqrt{K} (2K-5/2) 2^{K-1}} \; .
\end{align*}

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
$\norm{u_1'}^2 + \norm{u_2'}^2 + \dots + \norm{u_K'}^2 \le 1$.
Futhermore, for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma_1$ and
$\ip{u'_{y_t}}{x_t} \ge \gamma_2$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma_1$ and $\ip{u'_j}{x_t} \le - \gamma_2$. In other words,
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_1$ and also strongly linearly
separable with margin $\gamma_2$. Therefore, the examples are strongly
linearly separable with margin $\gamma' = \max\{\gamma_1, \gamma_2\}$.
\end{proof}
